---
title: "Covid91 vaccine study Final2023"
author: "Rita Jarbanda"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor)) 

male = filter(FinalData, sex =="M")
female = filter(FinalData, sex =="F")
lgbtq = filter(FinalData, LGBTQ =="gay")
drugUser = filter(FinalData, DrugUser =="yes")

```


# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

$H_A$ : There are statistically significant difference in the data, meaning that the vaccine and placebo are NOT equally as effective.

$H_0$ : There are no statistically significant difference in the data, meaning that the vaccine and placebo are equally as effective.

# Methods
Looking at the data collected, this will be categorical explained by categorical statistical problem, we will create separate data variables for each group to separate and test each of them. We will use the candP helper file to use R studios built in bar charts to display data graphically. W will also use xtabs to find the percentages relevant to the data. In this case we are only comparing two types of data and therefore we do not need to use any loops to measure multiple variables in the Fisher test or run a FisherXY.

## Males

### Graphical results

```{r}
barchartGC(~infected + treatment,data=male)
barchartGC(~infected + treatment,data=male, type="percent")
```

The above is a barchart that compares the graphical results of only males collected from the data. It is comparing rates of getting Covid 19 as compared by vaccination status. The left shows people who contracted Covid 19 and the right shows men who did not get the covid.

We can clearly see that more than twice the percentage of people who are in the placebo group got covid-19 where people who did not get covid are very close. Considering the sample size.

### Numerical Results

```{r}
tableMale <- xtabs(~infected + treatment, data=male)
rowPerc(tableMale)
colPerc(tableMale)
```

The numerical results is telling us that the placebo group contracted Covid are about 70% of the time where the vaccinated group contracted Covid 30% of the time. We can see that the vaccine is generally effective in this scenario.

### Inferential Results

```{r}
chisq.test(tableMale)
chisqtestGC(tableMale)
maleFisher <- fisher.test(tableMale)
maleFisher
```

The Chi Squared test shows that the P-value for the tested data is 0.00000000000004631 which smaller than the general use point of 0.05. This number indicates that there is a difference in the data. The Chi-Squared statistic is 55.18 which is way larger than our degrees of freedom of 1 in this data. 


### Confidence Interval

```{r}
oddsMale = maleFisher$estimate/(maleFisher$estimate+1)*100
print(oddsMale)
maleLow=(maleFisher$conf.int[1]/(maleFisher$conf.int[1]+1))*100
maleHigh=(maleFisher$conf.int[2]/(maleFisher$conf.int[2]+1))*100
cat("\n","Condifence interval for Males: ",maleLow," to ",maleHigh)
```

We see here that the confidence interval is more than 50% with a normal vaccine approach of more than 60%. The confidence interval shows us that this vaccine is good for this group.

## Females

### Graphical results

```{r}
barchartGC(~infected + treatment,data=female)
barchartGC(~infected + treatment,data=female, type="percent")
```
This is a barchart that compares the graphical results of famales only from the data. It is comparing rates of getting Covid 19 as compared by vaccination status. The left shows females who contracted Covid 19 and the right shows females who did not get covid.We can see that one third more of the percentage of people who are in the placebo group got covid while people who did not contract covid are very close. 

### Numerical Results

```{r}
tableFemale <- xtabs(~infected + treatment, data=female)
rowPerc(tableFemale)
colPerc(tableFemale)
```

The numerical results is telling us that the placebo group who got Covid are about 60% of the time where the vaccinated group got Covid 40% of the time. We can see that the vaccine is effective in this scenario however it was more effective of male participants.

### Inferential Results

```{r}
chisq.test(tableFemale)
chisqtestGC(tableFemale)
femaleFisher <- fisher.test(tableFemale)
femaleFisher
```

The Chi Squared test shows that the P-value for the tested data is 0.0002162 which smaller than 0.05. This number indicates that there is a difference in the data. The Chi-Squared statistic is 13.246  which is way larger than our degrees of freedom of 1 in this data. 

### Confidence Interval

```{r}
oddsFemale = femaleFisher$estimate/(femaleFisher$estimate +1)*100
print(oddsFemale)
femaleLow=(femaleFisher$conf.int[1]/(femaleFisher$conf.int[1]+1))*100
femaleHigh=(femaleFisher$conf.int[2]/(femaleFisher$conf.int[2]+1))*100
cat("\n","Condifence interval for Females: ",femaleLow," to ",femaleHigh)
```

We see here that the confidence interval is more than 50% with a normal vaccine approach of a percent that actually includes 60%. The confidence interval shows us that this vaccine is is less effective for females than males. 

## LGBTQ

### Graphical results

```{r}
barchartGC(~infected + treatment,data=lgbtq)
barchartGC(~infected + treatment,data=lgbtq, type="percent")
```

This is a barchart that compares the graphical results of LGBTQs only from the data. It is comparing rates of getting Covid 19 as compared by vaccination status. The left shows individuals who got Covid 19 and the right shows individuals who did not get covid.We can see that more vaccinated individuals got covid while people who did not get covid are very close. 

### Numerical Results

```{r}
tableLGBTQ <- xtabs(~infected + treatment, data=lgbtq)
rowPerc(tableLGBTQ)
colPerc(tableLGBTQ)
```

The numerical results is telling us that the placebo group who got Covid are about 15% of the time where the vaccinated group got Covid 85% of the time. We can see that the vaccine is not effective in this scenario.

### Inferential Results

```{r}
chisq.test(tableLGBTQ)
chisqtestGC(tableLGBTQ)
lgbtqFisher <- fisher.test(tableLGBTQ)
lgbtqFisher
```

The Chi Squared test shows that the P-value for the tested data is 0.00000001029 which smaller than 0.05. This number indicates that there is a difference in the data. The Chi-Squared statistic is 29.1588  which is way larger than our degrees of freedom of 1 in this data. 

### Confidence Interval

```{r}
oddsLGBTQ = lgbtqFisher$estimate/(lgbtqFisher$estimate +1)*100
print(oddsLGBTQ)
lgbtqLow=(lgbtqFisher$conf.int[1]/(lgbtqFisher$conf.int[1]+1))*100
lgbtqHigh=(lgbtqFisher$conf.int[2]/(lgbtqFisher$conf.int[2]+1))*100
cat("\n","Condifence interval for LGBTQ: ",lgbtqLow," to ",lgbtqHigh)
```

We see here that the confidence interval is much lower that 50% with a normal vaccine approach of a percent that that is also much lower than 60%. We can conclude that this vanccine in not good.

## Druggies


# Overall Results and Conclusions